Generalized canonical rings of $\bq$-divisors on minimal rational surfaces

### 2016-08-07 surfaces-write-up/MRL/MRL-surfaces-responses-to-referee.tex

\def\listtodoname{List of Todos}
\def\listoftodos{\@starttoc{tdo}\listtodoname}

\title{Summary of Changes in response to Referee's Report}
\author{Aaron Landesman, Peter Ruhm, and Robin Zhang}

\begin{document}

\maketitle
\section{Changes Made}
Here are the changes we have made in response to the referee's suggestions.
\begin{enumerate}
	\item We changed the name ``canonical ring'' to ``generalized canonical ring,'' as the referee suggested,
		since the canonical ring is really
		$R_K$ (i.e., it is the generalized canonical ring $R_D$ as we have defined it in the case $D = K$).
	\item We explained
		in more detail how our potential application
		for Hilbert modular forms and the difficulty in producing
		an explicit example.
	\item In all of the main theorems, it says that $\frac{c_i}{k_i}$ is in reduced form. 
	\item In Theorem \autoref{thm:proj-generators-relations}, $k$ is a field. Throughout our paper, the results hold for arbitrary fields, but we may assume the field is algebraically closed because the generator and relation degrees are preserved by base change to the algebraic closure. We have now included and clarified this in the ``Notations and Conventions'' subsection at the end of the Introduction.
	\item On page 2, line -9 of the previous version, $\deg(D)$ should have been $\frac{1}{2 \prod_{i=0}^n q_i}$ and is now corrected.
	\item Theorem 1.4 (now \autoref{thm:hirz-generators-relations}) has been rewritten into a simpler form by reducing some
		notation unnecessary to understanding the general idea. We then state it more precisely as \autoref{thm:hirz-generators-relations-full}. The statement of \autoref{thm:hirz-generators-relations-full} has also been simplified, since we noticed the previously defined quantity $\tau$ was unnecessary.
	\item Subsection 1.2 has now been moved to immediately precede Subsection 1.1 as a standalone paragraph.
	\item The reference to later items given on page 5, line 7 of the previous version
		has been moved to Remark \autoref{rem:proj-grobner}.
\item The use of best lower approximations has been replaced by the use of
convergents of Hirzebruch-Jung continued fractions which are equivalent and
well-known.
\item The references for the O'Dorney and Landesman--Ruhm--Zhang papers have been updated. The citation for arxiv articles now includes the arxiv reference numbers.
\item The conditions of Proposition~\autoref{prop:pm-span-and-basis} have been reorganized for clarity.
\item Remark 1.3 has been split into what is now Remark~\autoref{rem:proj-tight-bounds-eff} and Remark~\autoref{rem:proj-tight-bounds} (with
some material moved to Remark~\autoref{rem:exact-noneff-bounds}).

\end{enumerate}
\end{document}

### 2016-08-07 surfaces-write-up/surface-paper.tex
made canonical ring to generalized canonical ring

\title{Canonical rings of $\bq$-divisors on minimal rational surfaces}

@@

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{generalized canonical ring} is $R(X, D) := \bigoplus_{d \geq 0} u^d
H^0(X, \lfloor dD \rfloor)$, where $u$ is a dummy variable to keep
track of degree.
We study the generators and relations of canonical
rings for projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$.
When $D$ is a general $\bq$-divisor on $\bp^m$ or
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a minimal presentation of the
canonical ring when $X = \bp^m$ and $D$ is any effective
$\bq$-divisor.

In part, this work is motivated by a recent paper by
O'Dorney \cite{dorney:canonical}, which gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. This
study also has connections to the papers by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring},
which provide tight bounds on the degree of generators and relations
of log canonical rings and log spin canonical rings on arbitrary
stacky curves.

@@

\subsection{Notation}
We now define some notation used throughout the paper.
Use $m \in \bz$ to index the dimension of a given projective
space $\bp^m$ and the type of the Hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \bq \otimes_\bz \di X.
\end{align*}\noindent
where $n \in \bz$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is a divisor $\lfloor D
\rfloor \in \di X$ is defined to be $\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$. Let 
$R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)$
denote the canonical ring associated to the $\bq$-divisor $D$. 
So, $R_D$ is notation for $R(X,D)$ when $X$ is understood from context.
If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. 
Throughout, we work over a fixed algebraically closed field $\bk$.
Since generator and relation degrees are preserved under base change to the algebraic closure, assuming that $\bk$ is algebraically closed is no restriction
on generality and the results hold equally well over an arbitrary field.
If $D \in \bq \otimes_\bz
\di X$ is an arbitrary divisor, we denote $h^0(X, D) := \dim_\bk H^0(X,D)$.

@@

In Section \ref{sec:proj} we prove the following two results bounding
the degree of generators and relations of
stacky canonical rings on $\bp^m$. The first applies
to effective $\bq$-divisors and the second applies to arbitrary
$\bq$-divisors.
\begin{restatable}{thm}{restateeffective}
\label{thm:proj-effective-intro}
Let $D = \sum_{i = 0}^{n} \alpha_i D_i \in \bq \otimes_\bz \di \bp^m$, with $\alpha_i =
\frac{c_i}{k_i} \in \bq_{> 0}$ in reduced form and each $D_i \in \di \bp^m$ an integral divisor.

Then the canonical ring $R_D$
is generated in degrees at most $\max_{0 \leq i \leq n}{k_i}$ with
relations generated in degrees at most $2 \max_{0 \leq i \leq n}{k_i}$.
\end{restatable}

@@

\noindent
By the classification of minimal rational surfaces
as either $\bp^2$ or a Hirzebruch surface \cite{eisenbud-harris:minimal},
Theorems \ref{thm:proj-generators-relations} and
\ref{thm:hirz-generators-relations} provide bounds on generators
and relations for the canonical ring of any $\bq$-divisor on any
minimal rational surface. Section \ref{sec:conc} discusses further questions.

\section{Canonical Rings of Projective Space}
\label{sec:proj}
Let $\bk$ be a field and let $\bp^m$ denote $m$-dimensional
projective space over $\bk$. In this section, we bound the degrees
of generators and relations for divisors
on $\bp^m$, for all $m \geq 1$ in Theorem
~\ref{thm:proj-effective-intro}.
Note that if $\deg D < 0$ the
canonical ring is concentrated in degree 0, and if
$\deg D = 0$, then the canonical ring has a single
generator. Therefore, for the remainder of this section,
we shall assume $\deg D > 0$.
The $\bp^1$ case restricts to the results of
\cite{dorney:canonical}. We also give an explicit description of the
generators of the canonical ring $R_D$ when $D$ is an effective divisor in Theorem ~\ref{thm:proj-generators-relations}

@@

\ssec{Preliminaries on Projective Space}
Before bounding the degree of generation and relations of
canonical rings on projective space, we first
describe a particular choice of basis for $R_D$ as a $\bk$ vector
space.

@@

In this subsection, we restrict attention to the case of effective
fractional divisors $D \in \bq \otimes_\bz \di \bp^m$. We give an explicit presentation of the canonical ring when $D$ is an effective
divisor.

@@

\noindent
be the convergents of the Hirzebruch-Jung continued fraction of
$\alpha$. Then, the canonical ring
\[
	R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)
\]

@@

\begin{rem}
\label{rem:proj-two-points}
Note that we cannot extend this result to the case when
$D$ is supported at two hypersurfaces with arbitrary rational (non-effective)
coefficients in the same manner that O'Dorney does for the
$\bp^1$ case \cite[Section 4]{dorney:canonical}. As will be shown
in Example ~\ref{eg:hyperplane}, the degrees of generation of the
canonical ring of a general $\bq$-divisor supported on two
hyperplanes cannot be bounded so tightly. The two-point
$\bp^1$ result leverages the fact that $\bp^1$ has precisely two
independent coordinates, so that two distinct integral subschemes
cannot represent equations of only $m$ of the $m+1$ coordinates.
\end{rem}

Theorem ~\ref{thm:proj-one-point} gives an inductive procedure to
compute presentations of effective divisors which are supported on
hyperplanes. However, for $m \geq 2$, there are hypersurfaces which
are not unions of hyperplanes. We now address this general case, giving an inductive presentation of canonical rings of effective divisors, and a tight bound on the degrees of their generators and
relations.

@@

If $C$ is of degree $\pdeg$, consider the Veronese embedding
$\nu_\pdeg^m \colon \bp^m \rightarrow \bp^{\binom{m+\pdeg}{\pdeg}-1}$
so that the image of $C$ is the intersection of 
a hyperplane in $\bp^{\binom{{m + \pdeg}}{\pdeg} - 1}$ with $\nu_\pdeg^m(\bp^m)$.
Now, note that the ring $R_{\alpha C}$ is isomorphic to the
$\pdeg$ Veronese subring of $R_{\alpha V(x_0)} = \bigoplus_{d \geq 0} u^d H^0(\bp^m, d \alpha V(x_0))$ and so
the canonical ring $R_H$ of the hyperplane $H \subset \bp^{\binom{m+\pdeg}{\pdeg}-1}$
restricted to $\nu_\pdeg^m(C)$ is isomorphic to $R_C$.
Therefore, we can bound the degree of generators and relations of
$R_D$ over $R_{D'}$ by the degree of generators and
relations for $R_C \cong R_H.$
This reduces the case of a hypersurface $C$ to
a hyperplane $H$, completing the proof by Theorem
~\ref{thm:proj-one-point}.
\end{proof}

@@

By combining the above results, we get our main theorem bounding
the generator and relation degrees of the canonical ring of any
$\bq$-divisor on projective space.

@@

\section{Canonical Rings of Hirzebruch surfaces}
\label{sec:hirz}
The aim of this section is to prove Theorem
~\ref{thm:hirz-generators-relations}, bounding the degree of
generators and relations of the canonical ring of a $\bq$-divisor
on any Hirzebruch surface.

@@

Furthermore, similarly to the case of projective space, 
we will restrict to the case that $D$ is a divisor
for which both of its bi-degrees are positive. We now justify this
restriction. If either of the bi-degrees of $D$ are negative,
then the canonical ring is concentrated in degree 0. If one of the
bi-degrees is 0, say the first one is 0, then $R_D$ is isomorphic
to $R_{D'}$, where $D' \in \bq \otimes_\bz \di \bp^1$, 
where $D'$ can be written as a sum of divisors whose degrees
are multiples of the second bi-degree of $D$. Since the case of
$\bp^1$ has already been analyzed in \cite{dorney:canonical}, we
are justified in assuming that both bi-degrees of $D$ are positive.

@@

\begin{lem}
\label{lem:hirz-generators}
For $D = \sum_{i=1}^{n} \frac{c_i}{k_i}D_i \in \di \bq \otimes_\bz \hirz_m$, with $\deg D_i = (\bida_i, \bidb_i), \ell_i = \lcm_{j\neq i} (k_j), \ell_{i,j} = \lcm_{h \neq i,j} (k_h).$ Then, the canonical ring $R_D$ is generated in degrees at most
\begin{equation}\label{eqn:def-sigma}
	\rho := \sum_{i \in \Te(D)} \gcd(\bida_i, \bidb_i) \ell_i + \sum_{\substack
	{i \in \Tp(D) \\ j\in \Tm(D)}} (\bida_i \bidb_j - \bida_j \bidb_i)
	\ell	_{i,j}.
\end{equation}
\end{lem}

@@

\section{Further Questions}
\label{sec:conc}
Recall that every minimal rational surface is either isomorphic to
$\bp^2$ or $\hirz_m$ for some $m \geq 0, m \neq 1$ \cite{eisenbud-harris:minimal}. By Theorems
~\ref{thm:proj-generators-relations} and ~\ref{thm:hirz-generators-relations}, we have given bounds for the generators and relations
of arbitrary canonical rings on any minimal rational surface. A
natural extension of our results is the following.
\begin{question}
\label{qn:general-minimal-surface}
Can we describe generators and relations of $R_D$ for a divisor $D$ on an
arbitrary rational surface $X$?
\end{question}

Every rational surface can be obtained from a minimal rational surface by a 
sequence of blow-ups \cite{eisenbud-harris:minimal}. Therefore, to answer
Question \ref{qn:general-minimal-surface} affirmatively, it suffices
to bound the degree of generators and relations of the 
canonical ring of a divisor on a blow-up of a given surface
in terms of the canonical
rings of some associated divisors that given surface.

Another direction to generalize the work in this paper
would be to try to express canonical rings of $\bq$-divisors on
$X \times Y$ in terms of those on $X$ and $Y$. In this paper,
we bounded the degrees of presentations on canonical rings on $\bp^1 \times \bp^1 \cong \hirz_0$.
Perhaps similar techniques can be used to bound
degrees of presentations on canonical rings on $(\bp^1)^k$ or
more generally on $(\bp^1)^{i_1} \times \cdots (\bp^k)^{i_k}.$
One might further try to generalize this to bounding degrees of presentations on bundles over $\bp^m$ or on more general
products of schemes.

>>>>

\title{Generalized canonical rings of $\bq$-divisors on minimal rational surfaces}

@@

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{generalized canonical ring} is $R(X, D) := \bigoplus_{d \geq 0} u^d
H^0(X, \lfloor dD \rfloor)$, where $u$ is a dummy variable to keep
track of degree.
We study the generators and relations of generalized canonical
rings for projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$.
When $D$ is a general $\bq$-divisor on $\bp^m$ or
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a minimal presentation of the
generalized canonical ring when $X = \bp^m$ and $D$ is any effective
$\bq$-divisor.

In part, this work is motivated by a recent paper by
O'Dorney \cite{dorney:canonical}, which gives similar descriptions
of generalized canonical rings for general $\bq$-divisors on $\bp^1$. This
study also has connections to the papers by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring},
which provide tight bounds on the degree of generators and relations
of log canonical rings and log spin canonical rings on arbitrary
stacky curves.

@@

\subsection{Notation}
We now define some notation used throughout the paper.
Use $m \in \bz$ to index the dimension of a given projective
space $\bp^m$ and the type of the Hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \bq \otimes_\bz \di X.
\end{align*}\noindent
where $n \in \bz$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is a divisor $\lfloor D
\rfloor \in \di X$ is defined to be $\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$. Let 
$R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)$
denote the generalized canonical ring associated to the $\bq$-divisor $D$. 
So, $R_D$ is notation for $R(X,D)$ when $X$ is understood from context.
If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. 
Throughout, we work over a fixed algebraically closed field $\bk$.
Since generator and relation degrees are preserved under base change to the algebraic closure, assuming that $\bk$ is algebraically closed is no restriction
on generality and the results hold equally well over an arbitrary field.
If $D \in \bq \otimes_\bz
\di X$ is an arbitrary divisor, we denote $h^0(X, D) := \dim_\bk H^0(X,D)$.

@@

In Section \ref{sec:proj} we prove the following two results bounding
the degree of generators and relations of
generalized canonical rings on $\bp^m$. The first applies
to effective $\bq$-divisors and the second applies to arbitrary
$\bq$-divisors.
\begin{restatable}{thm}{restateeffective}
\label{thm:proj-effective-intro}
Let $D = \sum_{i = 0}^{n} \alpha_i D_i \in \bq \otimes_\bz \di \bp^m$, with $\alpha_i =
\frac{c_i}{k_i} \in \bq_{> 0}$ in reduced form and each $D_i \in \di \bp^m$ an integral divisor.

Then the generalized canonical ring $R_D$
is generated in degrees at most $\max_{0 \leq i \leq n}{k_i}$ with
relations generated in degrees at most $2 \max_{0 \leq i \leq n}{k_i}$.
\end{restatable}

@@

\noindent
By the classification of minimal rational surfaces
as either $\bp^2$ or a Hirzebruch surface \cite{eisenbud-harris:minimal},
Theorems \ref{thm:proj-generators-relations} and
\ref{thm:hirz-generators-relations} provide bounds on generators
and relations for the generalized canonical ring of any $\bq$-divisor on any
minimal rational surface. Section \ref{sec:conc} discusses further questions.

\section{Canonical Rings of Projective Space}
\label{sec:proj}
Let $\bk$ be a field and let $\bp^m$ denote $m$-dimensional
projective space over $\bk$. In this section, we bound the degrees
of generators and relations for divisors
on $\bp^m$, for all $m \geq 1$ in Theorem
~\ref{thm:proj-effective-intro}.
Note that if $\deg D < 0$ the
generalized canonical ring is concentrated in degree 0, and if
$\deg D = 0$, then the generalized canonical ring has a single
generator. Therefore, for the remainder of this section,
we shall assume $\deg D > 0$.
The $\bp^1$ case restricts to the results of
\cite{dorney:canonical}. We also give an explicit description of the
generators of the generalized canonical ring $R_D$ when $D$ is an effective divisor in Theorem ~\ref{thm:proj-generators-relations}

@@

\ssec{Preliminaries on Projective Space}
Before bounding the degree of generation and relations of
generalized canonical rings on projective space, we first
describe a particular choice of basis for $R_D$ as a $\bk$ vector
space.

@@

In this subsection, we restrict attention to the case of effective
fractional divisors $D \in \bq \otimes_\bz \di \bp^m$. We give an explicit presentation of the generalized canonical ring when $D$ is an effective
divisor.

@@

\noindent
be the convergents of the Hirzebruch-Jung continued fraction of
$\alpha$. Then, the generalized canonical ring
\[
	R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)
\]

@@

\begin{rem}
\label{rem:proj-two-points}
Note that we cannot extend this result to the case when
$D$ is supported at two hypersurfaces with arbitrary rational (non-effective)
coefficients in the same manner that O'Dorney does for the
$\bp^1$ case \cite[Section 4]{dorney:canonical}. As will be shown
in Example ~\ref{eg:hyperplane}, the degrees of generation of the
generalized canonical ring of a general $\bq$-divisor supported on two
hyperplanes cannot be bounded so tightly. The two-point
$\bp^1$ result leverages the fact that $\bp^1$ has precisely two
independent coordinates, so that two distinct integral subschemes
cannot represent equations of only $m$ of the $m+1$ coordinates.
\end{rem}

Theorem ~\ref{thm:proj-one-point} gives an inductive procedure to
compute presentations of effective divisors which are supported on
hyperplanes. However, for $m \geq 2$, there are hypersurfaces which
are not unions of hyperplanes. We now address this general case, giving an inductive presentation of generalized canonical rings of effective divisors, and a tight bound on the degrees of their generators and
relations.

@@

If $C$ is of degree $\pdeg$, consider the Veronese embedding
$\nu_\pdeg^m \colon \bp^m \rightarrow \bp^{\binom{m+\pdeg}{\pdeg}-1}$
so that the image of $C$ is the intersection of 
a hyperplane in $\bp^{\binom{{m + \pdeg}}{\pdeg} - 1}$ with $\nu_\pdeg^m(\bp^m)$.
Now, note that the ring $R_{\alpha C}$ is isomorphic to the
$\pdeg$ Veronese subring of $R_{\alpha V(x_0)} = \bigoplus_{d \geq 0} u^d H^0(\bp^m, d \alpha V(x_0))$ and so
the generalized canonical ring $R_H$ of the hyperplane $H \subset \bp^{\binom{m+\pdeg}{\pdeg}-1}$
restricted to $\nu_\pdeg^m(C)$ is isomorphic to $R_C$.
Therefore, we can bound the degree of generators and relations of
$R_D$ over $R_{D'}$ by the degree of generators and
relations for $R_C \cong R_H.$
This reduces the case of a hypersurface $C$ to
a hyperplane $H$, completing the proof by Theorem
~\ref{thm:proj-one-point}.
\end{proof}

@@

By combining the above results, we get our main theorem bounding
the generator and relation degrees of the generalized canonical ring of any
$\bq$-divisor on projective space.

@@

\section{Canonical Rings of Hirzebruch surfaces}
\label{sec:hirz}
The aim of this section is to prove Theorem
~\ref{thm:hirz-generators-relations}, bounding the degree of
generators and relations of the generalized canonical ring of a $\bq$-divisor
on any Hirzebruch surface.

@@

Furthermore, similarly to the case of projective space, 
we will restrict to the case that $D$ is a divisor
for which both of its bi-degrees are positive. We now justify this
restriction. If either of the bi-degrees of $D$ are negative,
then the generalized canonical ring is concentrated in degree 0. If one of the
bi-degrees is 0, say the first one is 0, then $R_D$ is isomorphic
to $R_{D'}$, where $D' \in \bq \otimes_\bz \di \bp^1$, 
where $D'$ can be written as a sum of divisors whose degrees
are multiples of the second bi-degree of $D$. Since the case of
$\bp^1$ has already been analyzed in \cite{dorney:canonical}, we
are justified in assuming that both bi-degrees of $D$ are positive.

@@

\begin{lem}
\label{lem:hirz-generators}
For $D = \sum_{i=1}^{n} \frac{c_i}{k_i}D_i \in \di \bq \otimes_\bz \hirz_m$, with $\deg D_i = (\bida_i, \bidb_i), \ell_i = \lcm_{j\neq i} (k_j), \ell_{i,j} = \lcm_{h \neq i,j} (k_h).$ Then, the generalized canonical ring $R_D$ is generated in degrees at most
\begin{equation}\label{eqn:def-sigma}
	\rho := \sum_{i \in \Te(D)} \gcd(\bida_i, \bidb_i) \ell_i + \sum_{\substack
	{i \in \Tp(D) \\ j\in \Tm(D)}} (\bida_i \bidb_j - \bida_j \bidb_i)
	\ell	_{i,j}.
\end{equation}
\end{lem}

@@

\section{Further Questions}
\label{sec:conc}
Recall that every minimal rational surface is either isomorphic to
$\bp^2$ or $\hirz_m$ for some $m \geq 0, m \neq 1$ \cite{eisenbud-harris:minimal}. By Theorems
~\ref{thm:proj-generators-relations} and ~\ref{thm:hirz-generators-relations}, we have given bounds for the generators and relations
of arbitrary generalized canonical rings on any minimal rational surface. A
natural extension of our results is the following.
\begin{question}
\label{qn:general-minimal-surface}
Can we describe generators and relations of $R_D$ for a divisor $D$ on an
arbitrary rational surface $X$?
\end{question}

Every rational surface can be obtained from a minimal rational surface by a 
sequence of blow-ups \cite{eisenbud-harris:minimal}. Therefore, to answer
Question \ref{qn:general-minimal-surface} affirmatively, it suffices
to bound the degree of generators and relations of the 
generalized canonical ring of a divisor on a blow-up of a given surface
in terms of the generalized canonical
rings of some associated divisors that given surface.

Another direction to generalize the work in this paper
would be to try to express generalized canonical rings of $\bq$-divisors on
$X \times Y$ in terms of those on $X$ and $Y$. In this paper,
we bounded the degrees of presentations on generalized canonical rings on $\bp^1 \times \bp^1 \cong \hirz_0$.
Perhaps similar techniques can be used to bound
degrees of presentations on generalized canonical rings on $(\bp^1)^k$ or
more generally on $(\bp^1)^{i_1} \times \cdots (\bp^k)^{i_k}.$
One might further try to generalize this to bounding degrees of presentations on bundles over $\bp^m$ or on more general
products of schemes.
